\vspace{-0.5in}{\bfseries\scshape Transfer Learning for computationally limited training of Vision Transformers on art classification tasks } \\ \vspace{1ex} \normalsize{Bachelor's Project Thesis}\vspace{-2ex}

### 2022-07-30 writing/report.tex
Made all the changes that I wrote down on the printed out draft. Up next: throwing it through Docs to check grammar, and having it read by one or two friends!
Made changes suggested by Grammarly

\twocolumn[
  \maketitle
  \begin{@twocolumnfalse}
   \input{tex/draft_abstract.tex}
  \end{@twocolumnfalse}
]

@@

\input{tex/draft_introduction.tex}
\input{tex/draft_methods.tex}
\input{tex/draft_experiments.tex}
\input{tex/draft_conclusion.tex}

\section*{Acknowledgements}
\textit{The author would like to thank {the Center for Information Technology of the University of Groningen} for their support and for providing access to the Peregrine high performance computing cluster. He would also like to thank {dr. M. Sabatelli} for his guidance during this project.}

\clearpage

>>>>

\twocolumn[
  \maketitle
  \begin{@twocolumnfalse}
   \input{tex/abstract.tex}
  \end{@twocolumnfalse}
]

@@

\input{tex/introduction.tex}
\input{tex/methods.tex}
\input{tex/experiments.tex}
\input{tex/conclusion.tex}

\section*{Acknowledgements}
\textit{The author would like to thank {the Center for Information Technology of the University of Groningen} for their support and for providing access to the Peregrine high-performance computing cluster. He would also like to thank {dr. M. Sabatelli} for his guidance during this project.}

### 2022-07-30 writing/tex/abstract.tex
Made changes suggested by Grammarly

\begin{abstract} {Convolutional Neural Networks (CNNs) have become the de facto standard in Computer Vision, and played a major role in the advancement of this field. In recent years, however, Vision Transformer based models (VTs) have been outperforming CNNs across the board. While this is exciting, CNNs still have an advantage on small datasets, due to their so-called image specific inductive bias. Recent work suggests that transfer learning methods allow VTs to compete with CNNs on some of these small datasets. This thesis investigates whether that also holds true for art classification problems within the digital humanities. To this end, it compares popular VTs and CNNs in terms of their off-the-shelf and fine-tuning transferability, when going from \textit{ImageNet1K} to target tasks provided by the \textit{Rijksmuseum Challenge} dataset. The results show that VTs posses superior off-the-shelf feature extraction capabilities here, and that in general, transfer learning allows VTs to become an interesting alternative to consider within the digital humanities.\vspace{8ex}}
\end{abstract}

>>>>

\begin{abstract} {Convolutional Neural Networks (CNNs) have become the de facto standard in Computer Vision, and played a major role in the advancement of this field. In recent years, however, Vision-Transformer-based models (VTs) have been outperforming CNNs across the board. While this is exciting, CNNs still have an advantage on small datasets, due to their so-called image-specific inductive bias. Recent work suggests that transfer learning methods allow VTs to compete with CNNs on some of these small datasets. This thesis investigates whether that also holds true for art classification problems within the digital humanities. To this end, it compares popular VTs and CNNs in terms of their off-the-shelf and fine-tuning transferability, when going from \textit{ImageNet1K} to target tasks provided by the \textit{Rijksmuseum Challenge} dataset. The results show that VTs possess superior off-the-shelf feature extraction capabilities here, and that in general, transfer learning allows VTs to become an interesting alternative to consider within the digital humanities.\vspace{8ex}}
\end{abstract}

### 2022-07-30 writing/tex/conclusion.tex
Made changes suggested by Grammarly

\section{Conclusion}
Seeing which is better, VTs or CNNs, has never been the goal of this paper. The underlying architecture is somewhat irrelevant, as one should always pick (or create) the model that fits the task best -- regardless of whether it is a VT, CNN, or something else. What this paper, then, was interested in, is whether VTs can become a valuable alternative to CNNs under circumstances in which:
\begin{itemize}
\item small datasets are used: specifically, small enough to have been hand labelled by one person;
\item only consumer grade hardware is available.
\end{itemize}
In other words: this paper was interested in the type of machine learning one could try at home, and wanted to know if VTs could be made useful for it.

To that end, it tried to utilise transfer learning methods, as these might help overcome problems that arise from VTs' lack of an inductive bias. More concretely, models pre-trained on ImageNet1K were either used as off-the-shelf feature extractors, or fine-tuned to new tasks. These tasks were all taken from the \textit{Rijksmuseum Challenge} dataset \citep{mensink14icmr}, and fall within the domain of art classification.

>>>>

\section{Conclusion}
Seeing which is better, VTs or CNNs, has never been the goal of this paper. The underlying architecture is somewhat irrelevant, as one should always pick (or create) the model that fits the task best -- regardless of whether it is a VT, CNN, or something else. What this paper, then, was interested in, is whether VTs can become a valuable alternative to CNNs under circumstances in which:
\begin{itemize}
\item small datasets are used: specifically, small enough to have been hand-labeled by one person;
\item only consumer-grade hardware is available.
\end{itemize}
In other words: this paper was interested in the type of machine learning one could try at home, and wanted to know if VTs could be made useful for it.

To that end, it tried to utilize transfer learning methods, as these might help overcome problems that arise from VTs' lack of an inductive bias. More concretely, models pre-trained on ImageNet1K were either used as off-the-shelf feature extractors, or fine-tuned to new tasks. These tasks were all taken from the \textit{Rijksmuseum Challenge} dataset \citep{mensink14icmr}, and fall within the domain of art classification.
